manuscript_figures

### 2024-07-06 scripts/manuscript_figures.Rmd
added production and HPI data

# Load libraries 
```{r}
library(plyr)
library(kableExtra)
library(vegan)
library(flextable)
library(gt)
library(dplyr)
library(tigris)
library(MASS)
library(htmlwidgets)
library(car)
library(ggiraphExtra)
library(MuMIn)
library(lmerTest)
library(tidyverse)
library(lme4)
library(broom.mixed)
library(ggmap)
library(lubridate)
library(cowplot)
library(ggplot2)
library(PerformanceAnalytics)
library(reshape2)
library(viridis)
library(GGally)
library(corrplot)
library(patchwork)
library(broom)
library(knitr)
library(ggspatial)
library(rslp)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggspatial)
library(gridExtra)
library(grid)

@@

# Assuming you want to join by the 'transect' column
merged.dat.cover <- left_join(fish.filt, kelp.filt)

@@

# Figure 3: organize and calculate kelp biomass data to make fish biomass/kelp biomass (g/m2) linear regression figure
```{r}
# read in quad data with filtered for just kelp
quad.filt <- QUAD %>%
        filter(SP_CODE == "SUGK")

@@

# Merge fish and morph data
merged.dat.morph <- left_join(morph.filt, fish.filt)

@@

# Fit the linear mixed effects model with log transformation
lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_kelp_biomass + 1.1) + SITE + (1 | SITE/YEAR), data = aggregated.data.morph)
#check normality
qqPlot(residuals(lmm_model)) #improved to normal

# Identify the outliers
outliers <- c(45, 37, 25, 33, 37, 22)

@@

# Plotting using ggplot2
production.plot <- ggplot(PB_filtered, aes(x = YEAR, y = PB.ratio, fill = SP_CODE)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Species",
       y = "P/B",
       fill = "Species") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),  # Rotate x-axis labels if needed
        axis.title.x = element_text(vjust = -0.5, size = 18),  # Adjust x-axis title position
        axis.title.y = element_text(vjust = 1.5, size = 18),
        axis.text.y = element_text(size = 18),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 18)) + # Adjust y-axis title position
  scale_y_continuous(limits = c(0, 4), expand = c(0, 0));production.plot  # Set y-axis limits and remove expansion

@@

# We add a small constant (1e-6) to avoid issues with log(0)
HPI_calculated <- HPI_calculated %>%
  mutate(log_HPI = log(HPI + 1.1))

# Now plot with the log-transformed HPI
species.hpi.plot <- ggplot(HPI_calculated, aes(x = SITE, y = log_HPI, fill = factor(YEAR))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Site", 
       y = "Log Habitat Productivity Index", 
       fill = "Year") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),
        axis.text.y = element_text(size = 18),
        axis.title.y = element_text(size = 18),
         axis.title.x = element_text(size = 18),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 18)) +
  scale_y_continuous(expand = expansion(mult = c(0, .05)),
    labels = scales::label_number(accuracy = 0.1)) +
  coord_cartesian(clip = "off"); species.hpi.plot

@@

# Plot two plots together using patchwork for production and HPI
```{r}
# Combine the plots with patchwork for all years of kelp and rate of change
production.hpi <- production.plot | species_hpi.plot +
  plot_annotation(tag_levels = 'A') + 
  theme(plot.tag = element_text(size = 20, face = "bold"));production.hpi

>>>>

# Load libraries 
```{r}
library(plyr)
library(kableExtra)
library(vegan)
library(RColorBrewer)
library(flextable)
library(gt)
library(dplyr)
library(tigris)
library(MASS)
library(htmlwidgets)
library(car)
library(ggiraphExtra)
library(MuMIn)
library(lmerTest)
library(tidyverse)
library(lme4)
library(broom.mixed)
library(ggmap)
library(lubridate)
library(cowplot)
library(ggplot2)
library(PerformanceAnalytics)
library(reshape2)
library(viridis)
library(GGally)
library(corrplot)
library(patchwork)
library(broom)
library(knitr)
library(ggspatial)
library(rslp)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggspatial)
library(gridExtra)
library(grid)

@@

# add 2018 fish data to fish.filt dataframe
fish.2018 <- read.csv("data/fish.data.clean.calc.csv")

# Filter for only 2018 data
fish.2018.filtered <- subset(fish.2018, YEAR == 2018)

# Rename kings beach and fort wetherill
fish.2018.filtered$SITE[fish.2018.filtered$SITE == "Kings Beach"] <- "King's Beach"
fish.2018.filtered$SITE[fish.2018.filtered$SITE == "Fort Weatherill"] <- "Fort Wetherill"

# Calculate fish biomass 'W' using the length-weight relationship
fish.2018.filtered$Wmat <- fish.2018.filtered$coef.a * (fish.2018.filtered$Lmat^fish.2018.filtered$coef.b)

# Calculate fish biomass
fish.2018.filtered$Fish.Biomass..g.80m2. <- fish.2018.filtered$Wmat * fish.2018.filtered$COUNT

# Calculate biomass per m²
fish.2018.filtered$Fish.Biomass..g.m2. <- fish.2018.filtered$Fish.Biomass..g.80m2. / 80

# selcet specific columns needed to merge
fish.2018.selected <- fish.2018.filtered[, c("YEAR", "HABITAT", "REGION", "SITE", "TRANSECT", "COUNT", "Fish.Biomass..g.m2.", "Fish.Biomass..g.80m2.")]

# Convert the integer column to character
fish.2018.selected$TRANSECT <- as.character(fish.2018.selected$TRANSECT)

#add area and n column
fish.2018.selected$AREA <- 80
fish.2018.selected <- fish.2018.selected %>%
  rename(n = COUNT)

# Merge (rbind) fish.2018.selected with fish.filt
fish.selected <- rbind(fish.filt, fish.2018.selected)

# Assuming you want to join by the 'transect' column
merged.dat.cover <- left_join(fish.selected, kelp.filt)

@@

# Figure 3: organize and calculate kelp biomass data to make fish biomass/kelp biomass (g/m2) linear regression figure
```{r}
# read in quad data with filtered for just kelp
quad.filt <- QUAD %>%
  filter(SP_CODE == "SUGK" | SP_CODE == "LADI")

@@

# Merge fish and morph data
merged.dat.morph <- left_join(morph.filt, fish.selected)

@@

# Fit the linear mixed effects model with log transformation
lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_kelp_biomass + 1.1) + SITE + (1 | SITE/YEAR), data = aggregated.data.morph)
#check normality
qqPlot(residuals(lmm_model)) #improved to normal
summary(lmm_model)

# Identify the outliers
outliers <- c(45, 37, 29,37)

@@

# Convert YEAR to factor (if it's numeric)
PB_filtered$YEAR <- factor(PB_filtered$YEAR)

# Calculate production by multiplying biomass_per_m2 by PB raatio
PB_filtered$Production <- PB_filtered$PB.ratio * PB_filtered$biomass_per_m2

# make commmon name column
PB_filtered <- PB_filtered %>%
  mutate(Species = case_when(
    SP_CODE == "BRTY" ~ "Atlantic menhaden",
    SP_CODE == "CEST" ~ "Black sea bass",
    SP_CODE == "MOSA" ~ "Striped bass",
    SP_CODE == "PADE" ~ "Summer flounder",
    SP_CODE == "STCH" ~ "Scup",
    SP_CODE == "TAAD" ~ "Cunner",
    SP_CODE == "TAON" ~ "Tautog",
    TRUE ~ SP_CODE  # This will keep the original code if it doesn't match any of the above
  ))

# Now you can use this new 'Species' column in your ggplot
production_plot <- ggplot(PB_filtered, aes(x = YEAR, y = Production, fill = Species)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Year",  # X-axis label
       y = expression("Fish Production" ~ (g ~ m^-2 ~ yr^-1)),  # Y-axis label
       fill = "Species") +  # This should now correctly set the legend title
  scale_fill_viridis_d() +  # Use viridis colorblind-friendly palette
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12),  # Rotate x-axis labels if needed
        axis.title.x = element_text(vjust = -0.5, size = 14),  # Adjust x-axis title position
        axis.title.y = element_text(vjust = 1.5, size = 14),
        axis.text.y = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14)) +
  scale_y_continuous(limits = c(0, 4), expand = c(0, 0));production_plot  # Additional attempt to change legend title  # Use viridis colorblind-friendly palette

@@

# Log transform HPI 
HPI_calculated <- HPI_calculated %>%
  mutate(log_HPI = log(HPI + 1.1))

# Define the color palette
my_palette <- brewer.pal(n = 6, name = "Paired")  # Adjust 'n' as per the number of years in your data

# Plotting code
hpi.plot <- ggplot(HPI_calculated, aes(x = SITE, y = log_HPI, fill = factor(YEAR))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Site", 
       y = "Log Habitat Productivity Index", 
       fill = "Year") +
  scale_fill_manual(values = my_palette) +  # Use the colorblind-friendly palette
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),
        axis.text.y = element_text(size = 18),
        axis.title.y = element_text(size = 18),
        axis.title.x = element_text(size = 18),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 18)) +
  scale_y_continuous(expand = expansion(mult = c(0, .05)),
                     labels = scales::label_number(accuracy = 0.1)) +
  coord_cartesian(clip = "off");hpi.plot

@@

# Plot two plots together using patchwork for production and HPI
```{r}
# Combine the plots with patchwork for all years of kelp and rate of change
production.hpi <- production_plot | hpi.plot +
  plot_annotation(tag_levels = 'A') + 
  theme(plot.tag = element_text(size = 20, face = "bold"));production.hpi

# Save the combined plot
ggsave("output/figures/production.HPI.png", production.hpi, width = 20, height = 10, units = "in", dpi = 300)
